---
title: "FiberMargin: Complete User Guide"
author: "FiberMargin authors"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{FiberMargin: Complete User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7.2,
  fig.height = 3.2,
  fig.align = "center",
  dpi = 110
)
options(width = 78)
library(fibermargin)

field_palette <- function(labels) {
  classes <- sort(unique(as.character(labels[!is.na(labels)])))
  stats::setNames(grDevices::hcl.colors(length(classes), "Dark 3"), classes)
}

plot_field <- function(xy, labels, main = NULL, cex = 0.55) {
  palette <- field_palette(labels)
  graphics::plot(
    xy[, 1L], xy[, 2L],
    col = palette[as.character(labels)],
    pch = 16, cex = cex, asp = 1,
    xlab = colnames(xy)[1L], ylab = colnames(xy)[2L],
    main = main
  )
}

plot_mask <- function(mask, main = NULL, palette = field_palette(mask)) {
  values <- matrix(
    match(as.character(mask), names(palette)),
    nrow = nrow(mask), ncol = ncol(mask)
  )
  graphics::image(
    seq_len(nrow(mask)), seq_len(ncol(mask)), values,
    col = unname(palette), useRaster = TRUE, axes = FALSE,
    xlab = "", ylab = "", main = main, asp = 1
  )
}
```

# Purpose and input contract

FiberMargin repairs a categorical field from spatial coordinates and the observed
labels alone. It does not use expression values, image intensities, logits,
probabilities, clean anchors, or reference labels during repair.

There are two fixed computational routes:

- Multiclass fields use the two-sided path-enclosure operator.
- Binary fields use a distinct fixed nearest-neighbour ballot specialization.

The package accepts irregular 2D or 3D coordinates through
`refine_spatial_labels()` and regular pixel or voxel arrays through
`clean_categorical_mask()`. Multiple specimens can be supplied in one call and
are always processed independently.

The package exports 14 functions:

```{r api-index}
api <- data.frame(
  Function = c(
    "refine_spatial_labels", "clean_categorical_mask",
    "corrupt_categorical_mask", "evaluate_mask_cleaning",
    "spatial_benchmark", "evaluate_spatial_refinement",
    "benchmark_spatial_refiners", "simulate_spatial_clusters",
    "simulate_spatial_domains", "simulate_complex_spatial_domains",
    "simulate_gradient_regions", "simulate_volumetric_domains",
    "available_spatial_benchmarks", "load_spatial_benchmark"
  ),
  Purpose = c(
    "Refine labels at irregular 2D/3D coordinates",
    "Clean a categorical matrix or 3D array",
    "Generate reproducible mask errors",
    "Score a cleaned mask",
    "Validate and package a benchmark",
    "Score coordinate-indexed refinement",
    "Run identical inputs through multiple methods",
    "Simulate separated Gaussian-like clusters",
    "Simulate structured 2D or curved-layer 3D domains",
    "Simulate held-out complex geometries",
    "Simulate the A-B-C gradient design",
    "Simulate variable-density 3D domains",
    "List real-data availability and licensing",
    "Load a bundled frozen real-data scenario"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(api)
```

# Quick start

The gradient simulator creates four ordered areas with reference classes
`A`, `B`, `B`, and `C`. Its observed labels contain controlled minority mixtures.

```{r quick-start}
gradient <- simulate_gradient_regions(
  n = 1200L,
  minority = 0.10,
  dimensions = 2L,
  samples = 2L,
  density_profile = "moderate",
  seed = 12L
)

refined_gradient <- refine_spatial_labels(
  xy = gradient$xy,
  labels = gradient$labels,
  samples = gradient$samples,
  workers = 1L
)

gradient_metrics <- evaluate_spatial_refinement(
  truth = gradient$truth,
  initial = gradient$labels,
  refined = refined_gradient,
  boundary = gradient$boundary,
  regions = gradient$area,
  sparse = gradient$sparse,
  method = "FiberMargin"
)

gradient_metrics[, c(
  "method", "initial_accuracy", "accuracy", "ari",
  "correction_recall", "damage_rate"
)]
```

```{r quick-start-plot, fig.height=3.0}
old_par <- graphics::par(mfrow = c(1, 3), mar = c(3, 3, 2, 1))
plot_field(gradient$xy, gradient$truth, "Reference", cex = 0.35)
plot_field(gradient$xy, gradient$labels, "Observed", cex = 0.35)
plot_field(gradient$xy, refined_gradient, "FiberMargin", cex = 0.35)
graphics::par(old_par)
```

# Refine irregular coordinates

## `refine_spatial_labels()`

```r
refine_spatial_labels(xy, labels, samples = NULL, workers = NULL)
```

`xy` must be a finite numeric matrix with two or three columns. `labels` must
contain one non-missing categorical value per row. `samples` is optional and
defines independent coordinate systems. Integer, character, and factor specimen
identifiers are accepted. `workers = NULL` uses a conservative automatic CPU
budget; set a positive integer for an explicit total budget.

Effective dimensionality is determined separately within every specimen.
Zero-range axes are removed before refinement. At least two varying axes must
remain, so constant `z` reduces a three-column specimen to 2D while variable
`z` remains part of a genuine 3D analysis.

The output is a factor with the input levels and row names. Seven pointwise audit
attributes are attached:

```{r diagnostics}
diagnostic_names <- c(
  "candidate", "margin_score", "required", "repair_margin",
  "atlas_dispersion", "isolation", "changed"
)

diagnostics <- data.frame(
  observed = head(as.character(gradient$labels), 8L),
  refined = head(as.character(refined_gradient), 8L),
  candidate = head(as.character(attr(refined_gradient, "candidate")), 8L),
  margin_score = head(attr(refined_gradient, "margin_score"), 8L),
  required = head(attr(refined_gradient, "required"), 8L),
  repair_margin = head(attr(refined_gradient, "repair_margin"), 8L),
  changed = head(attr(refined_gradient, "changed"), 8L)
)
diagnostics

setdiff(diagnostic_names, names(attributes(refined_gradient)))
```

`candidate` is the locally preferred class. `margin_score` is its support
contrast, `required` is the local admission barrier, and `repair_margin` is
`margin_score - required`. These are deterministic scores, not calibrated
probabilities. `atlas_dispersion` describes disagreement across path charts,
`isolation` protects locally isolated observations, and `changed` identifies
accepted edits.

Eight summary attributes provide a specimen-level audit:

```{r summary-diagnostics}
summary_names <- c(
  "workers", "dimensions_used", "labels_changed", "changed_fraction",
  "classes_before", "classes_after", "removed_classes", "sample_sizes"
)
lapply(summary_names, function(name) attr(refined_gradient, name))
```

`dimensions_used`, `labels_changed`, `changed_fraction`, and `sample_sizes` are
named vectors. The three class summaries are named lists. An unused factor level
is not counted as a class present before refinement. FiberMargin does not impose
class preservation; if an observed class disappears, it is listed in
`removed_classes`.

The output contract can be checked directly:

```{r diagnostic-contract}
identical(
  as.logical(attr(refined_gradient, "changed")),
  as.character(refined_gradient) != as.character(gradient$labels)
)
attr(refined_gradient, "workers")
```

## Binary specialization

With exactly two observed classes, the function uses the fixed local ballot
instead of the multiclass path enclosure.

```{r binary-specialization}
binary_data <- simulate_spatial_clusters(
  n = 700L,
  dimensions = 2L,
  k = 2L,
  samples = 1L,
  noise = 0.12,
  seed = 21L
)
binary_refined <- refine_spatial_labels(
  binary_data$xy, binary_data$labels,
  samples = binary_data$samples, workers = 1L
)

c(
  input_accuracy = mean(binary_data$labels == binary_data$truth),
  refined_accuracy = mean(binary_refined == binary_data$truth),
  changed_fraction = mean(binary_refined != binary_data$labels)
)
```

For this route, `isolation` is one and `atlas_dispersion` is zero because no
multiclass path atlas is used.

## Effective 2D and genuine 3D coordinates

A constant third axis is removed within the specimen, producing the same labels
and every same pointwise diagnostic as the 2D call.

```{r constant-z}
flat_2d <- refine_spatial_labels(
  gradient$xy, gradient$labels, gradient$samples, workers = 1L
)
flat_3d <- refine_spatial_labels(
  cbind(gradient$xy, z = 0), gradient$labels,
  gradient$samples, workers = 1L
)
identical(flat_2d, flat_3d)
attr(flat_3d, "dimensions_used")
```

For variable `z`, all three coordinates participate in chart construction and
distance calculations.

```{r genuine-3d}
volume_example <- simulate_volumetric_domains(
  n = 1800L,
  shape = "folded_layers",
  samples = 2L,
  noise = 0.18,
  seed = 23L
)
volume_refined <- refine_spatial_labels(
  volume_example$xy, volume_example$labels,
  volume_example$samples, workers = 2L
)
attr(volume_refined, "dimensions_used")
```

## Multiple specimens

Specimen identifiers prevent evidence from crossing tissue or section boundaries.
Numeric coordinates may overlap completely because each specimen has its own
coordinate system. The combined result is equivalent to refining each specimen
separately and is restored to the original row order.

```{r specimen-isolation}
source_rows <- which(gradient$samples == levels(gradient$samples)[1L])
overlapping_xy <- rbind(
  gradient$xy[source_rows, , drop = FALSE],
  gradient$xy[source_rows, , drop = FALSE]
)
overlapping_labels <- factor(
  rep(as.character(gradient$labels[source_rows]), 2L),
  levels = levels(gradient$labels)
)
overlapping_samples <- rep(
  c("section_1", "section_2"), each = length(source_rows)
)
joint <- refine_spatial_labels(
  overlapping_xy, overlapping_labels,
  samples = overlapping_samples, workers = 2L
)

separate <- unsplit(
  lapply(split(seq_len(nrow(overlapping_xy)), overlapping_samples), function(rows) {
    refine_spatial_labels(
      overlapping_xy[rows, , drop = FALSE],
      overlapping_labels[rows],
      workers = 1L
    )
  }),
  overlapping_samples
)

identical(as.character(joint), as.character(separate))
```

## Deterministic CPU use

`workers` is one total native budget, not workers per specimen. Specimens reuse
the budget and native stages do not create nested pools. The implementation uses
standard C++ threads on macOS, Linux, and Windows. Changing `workers` changes
execution only, not labels or diagnostics.

```{r worker-determinism, eval=FALSE}
one_worker <- refine_spatial_labels(
  gradient$xy, gradient$labels, gradient$samples, workers = 1L
)
four_workers <- refine_spatial_labels(
  gradient$xy, gradient$labels, gradient$samples, workers = 4L
)
attr(one_worker, "workers") <- NULL
attr(four_workers, "workers") <- NULL
identical(one_worker, four_workers)
```

# Clean categorical masks

## `corrupt_categorical_mask()`

This function creates reproducible test errors in a reference matrix or 3D array.
Available mechanisms are:

- `"impulse"`: independently scattered substitutions.
- `"boundary"`: substitutions concentrated near class boundaries.
- `"patch"`: one coherent overwritten spatial patch.

Every reference class retains at least one uncorrupted exemplar.

```{r construct-mask}
mask_size <- 48L
mask_grid <- expand.grid(row = seq_len(mask_size), column = seq_len(mask_size))
radius <- sqrt(
  (mask_grid$row - (mask_size + 1) / 2)^2 +
    (mask_grid$column - (mask_size + 1) / 2)^2
)

reference_mask <- matrix("background", mask_size, mask_size)
reference_mask[radius < 17] <- "ring"
reference_mask[radius < 10] <- "core"
reference_mask[1:4, 1:4] <- NA_character_

mask_corruptions <- lapply(
  c("impulse", "boundary", "patch"),
  function(mechanism) {
    corrupt_categorical_mask(
      reference_mask, mechanism = mechanism,
      rate = 0.15, seed = 100L
    )
  }
)
names(mask_corruptions) <- c("impulse", "boundary", "patch")

vapply(
  mask_corruptions,
  function(x) mean(x != reference_mask, na.rm = TRUE),
  numeric(1L)
)
```

## `clean_categorical_mask()`

```r
clean_categorical_mask(mask, samples = NULL, workers = NULL)
```

The function preserves dimensions, dimnames, storage type, and `NA` void cells.
It attaches diagnostic arrays with the same dimensions as the mask.

```{r clean-mask}
initial_mask <- mask_corruptions$boundary
cleaned_mask <- clean_categorical_mask(initial_mask, workers = 1L)

c(
  void_cells_preserved = all(is.na(cleaned_mask) == is.na(initial_mask)),
  changed_cells = sum(attr(cleaned_mask, "changed"), na.rm = TRUE)
)
dim(attr(cleaned_mask, "repair_margin"))
```

```{r mask-plot, fig.height=2.7}
mask_colors <- field_palette(reference_mask)
old_par <- graphics::par(mfrow = c(1, 3), mar = c(1, 1, 2, 1))
plot_mask(reference_mask, "Reference", mask_colors)
plot_mask(initial_mask, "Boundary corruption", mask_colors)
plot_mask(cleaned_mask, "FiberMargin", mask_colors)
graphics::par(old_par)
```

A sample array can isolate sections inside one mask:

```{r sampled-mask, eval=FALSE}
mask_samples <- matrix(
  ifelse(col(reference_mask) <= mask_size / 2, "left", "right"),
  nrow = mask_size, ncol = mask_size
)
cleaned_by_section <- clean_categorical_mask(
  initial_mask, samples = mask_samples, workers = 2L
)
```

The same API accepts a three-dimensional voxel array:

```{r voxel-mask, eval=FALSE}
voxel_reference <- array("A", dim = c(20, 20, 8))
voxel_reference[, 8:14, ] <- "B"
voxel_reference[, 15:20, ] <- "C"
voxel_initial <- corrupt_categorical_mask(
  voxel_reference, mechanism = "boundary", rate = 0.10, seed = 9L
)
voxel_cleaned <- clean_categorical_mask(voxel_initial, workers = 2L)
```

## `evaluate_mask_cleaning()`

Mask evaluation reports global recovery, ARI, class recall, correction and damage,
mean and worst IoU, Dice, boundary IoU, and rare-class IoU.

```{r evaluate-mask}
mask_metrics <- evaluate_mask_cleaning(
  reference = reference_mask,
  initial = initial_mask,
  cleaned = cleaned_mask,
  method = "FiberMargin"
)

mask_metrics[, c(
  "method", "initial_accuracy", "accuracy", "ari",
  "mean_iou", "mean_boundary_iou", "rare_class_iou",
  "correction_recall", "damage_rate"
)]
```

Accuracy obeys the exact repair-damage decomposition

```{r repair-damage-identity}
with(
  mask_metrics,
  c(
    reported = accuracy,
    decomposed = initial_accuracy * (1 - damage_rate) +
      (1 - initial_accuracy) * correction_recall,
    numerical_error = repair_identity_error
  )
)
```

# Build and evaluate benchmark objects

## `spatial_benchmark()`

Use `spatial_benchmark()` to validate custom coordinates and align all optional
evaluation strata. Reference labels are stored for evaluation and are not passed
to the repair function.

```{r spatial-benchmark}
custom_benchmark <- spatial_benchmark(
  xy = gradient$xy,
  labels = gradient$labels,
  truth = gradient$truth,
  samples = gradient$samples,
  boundary = gradient$boundary,
  regions = gradient$area,
  sparse = gradient$sparse,
  name = "two_gradient_tissues",
  metadata = list(seed = 12L, purpose = "vignette")
)

class(custom_benchmark)
names(custom_benchmark)
```

`boundary` and `sparse` are optional logical vectors. If `sparse` is omitted but
`regions` is supplied, evaluation uses the least frequent region as the sparse
stratum and the most frequent region as the dense stratum.

## `evaluate_spatial_refinement()`

This lower-level evaluator accepts aligned vectors and optional evaluation masks.

```{r evaluate-spatial}
manual_metrics <- evaluate_spatial_refinement(
  truth = custom_benchmark$truth,
  initial = custom_benchmark$labels,
  refined = refined_gradient,
  boundary = custom_benchmark$boundary,
  regions = custom_benchmark$regions,
  sparse = custom_benchmark$sparse,
  elapsed = 0,
  method = "FiberMargin"
)

manual_metrics[, c(
  "accuracy", "accuracy_gain", "ari", "macro_recall", "worst_recall",
  "boundary_accuracy", "sparse_region_accuracy",
  "correction_recall", "damage_rate", "changed_precision"
)]
```

Important measures are:

- `accuracy` and `ari`: global recovery.
- `correction_recall`: fraction of initially wrong labels repaired.
- `damage_rate`: fraction of initially correct labels made wrong.
- `changed_precision`: fraction of edited locations correct after repair.
- `macro_recall` and `worst_recall`: class-balance diagnostics.
- `boundary_accuracy` and `sparse_region_accuracy`: difficult-stratum outcomes.

## `benchmark_spatial_refiners()`

A method is any function that accepts `xy` and `labels`; it may also accept
`samples` or `...`. Every method receives identical input. The function accepts
one benchmark or a named list of benchmarks.

```{r benchmark-refiners}
jagged_small <- simulate_spatial_domains(
  n = 600L,
  pattern = "jagged_stripes",
  k = 4L,
  noise = 0.18,
  noise_type = "boundary",
  seed = 31L
)

comparison <- benchmark_spatial_refiners(
  data = list(
    gradient = custom_benchmark,
    jagged = jagged_small
  ),
  methods = list(
    FiberMargin = function(xy, labels, samples = NULL) {
      refine_spatial_labels(xy, labels, samples, workers = 1L)
    },
    IdentityExample = function(xy, labels, ...) labels
  ),
  include_initial = TRUE,
  seed = 5L,
  on_error = "stop"
)

comparison[, c(
  "dataset", "method", "accuracy", "ari",
  "correction_recall", "damage_rate", "seconds"
)]
```

Use `on_error = "record"` in long benchmark batches to retain a row containing
the error message instead of stopping the entire run.

# Simulation functions

Every simulator returns a `spatial_refinement_benchmark`-compatible list with
`xy`, noisy `labels`, reference `truth`, and `samples`. Most also include
`boundary`, `sparse`, region identifiers, and scenario metadata.

## `simulate_spatial_clusters()`

This is the simplest geometry: separated Gaussian-like clusters in two or three
dimensions. It is useful for smoke tests and binary-specialization examples.

```{r simulate-clusters}
cluster_sim <- simulate_spatial_clusters(
  n = 600L,
  dimensions = 3L,
  k = 4L,
  samples = 2L,
  noise = 0.10,
  seed = 2L
)
c(dimensions = ncol(cluster_sim$xy), classes = nlevels(cluster_sim$truth))
table(cluster_sim$samples)
```

## `simulate_spatial_domains()`

This simulator covers structured tissue geometries, corruption types, feature
widths, density profiles, and multiple specimens.

Available patterns are `jagged_stripes`, `wavy_layers`, `rings`, `spiral`,
`branching`, `lobes`, `islands`, `disconnected`, `thin_layers`, `intermixed`,
and `layers3d`. Only `layers3d` uses three dimensions. Corruption can be
`random`, `boundary`, `patch`, or `region`. Density can be `uniform`,
`moderate`, `strong`, `extreme`, `hotspot`, or a positive weight per class.

```{r simulate-domains}
domain_sim <- simulate_spatial_domains(
  n = 800L,
  pattern = "thin_layers",
  k = 5L,
  noise = 0.20,
  dimensions = 2L,
  samples = 2L,
  noise_type = "boundary",
  feature_scale = 0.8,
  density_profile = "strong",
  seed = 3L
)
table(domain_sim$truth)
```

## `simulate_complex_spatial_domains()`

This held-out geometry family stresses junctions, narrow channels, non-convex
boundaries, and severe class-density imbalance. Shapes are `voronoi_mosaic`,
`radial_sectors`, `checkerboard_junctions`, `tubular_network`,
`braided_channels`, and `shells_3d`.

```{r simulate-complex}
complex_sim <- simulate_complex_spatial_domains(
  n = 700L,
  shape = "tubular_network",
  density_profile = "extreme",
  noise_type = "patch",
  noise = 0.20,
  samples = 1L,
  k = 5L,
  seed = 4L
)
c(dimensions = ncol(complex_sim$xy), observations = nrow(complex_sim$xy))
```

## `simulate_gradient_regions()`

This function generated the quick-start object. It supports 2D or 3D coordinates,
multiple tissues, curved boundaries, and custom area-density weights.

```{r simulate-gradient-custom}
gradient_3d <- simulate_gradient_regions(
  n = 700L,
  minority = 0.05,
  dimensions = 3L,
  samples = 2L,
  seed = 5L,
  curvature = 0.06,
  density_profile = c(1, 2, 0.5, 3)
)
table(gradient_3d$area, gradient_3d$truth)
```

## `simulate_volumetric_domains()`

This is the dedicated 3D stress-test generator. Available shapes are
`concentric_shells`, `warped_ellipsoids`, `toroidal_compartments`,
`folded_layers`, `thin_folded_sheets`, `branching_tubes`,
`disconnected_volumes`, and `helical_channels`.

Acquisition can be `uniform`, `class_imbalanced`, or `irregular_z`; corruption
can be `random`, `boundary`, `patch`, or `region`.

```{r simulate-volume}
volume_sim <- simulate_volumetric_domains(
  n = 700L,
  shape = "folded_layers",
  acquisition = "class_imbalanced",
  noise_type = "boundary",
  noise = 0.20,
  k = 5L,
  samples = 2L,
  seed = 6L
)

c(
  dimensions = ncol(volume_sim$xy),
  observations = nrow(volume_sim$xy),
  samples = nlevels(volume_sim$samples)
)
volume_sim$region_counts
```

```{r simulator-gallery, fig.height=5.8}
old_par <- graphics::par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
plot_field(cluster_sim$xy, cluster_sim$truth, "Spatial clusters", cex = 0.45)
plot_field(domain_sim$xy, domain_sim$truth, "Thin layers", cex = 0.45)
plot_field(complex_sim$xy, complex_sim$truth, "Tubular network", cex = 0.45)
plot_field(volume_sim$xy, volume_sim$truth, "3D folded layers: x-y", cex = 0.45)
graphics::par(old_par)
```

# Bundled real benchmarks

## `available_spatial_benchmarks()`

Always inspect availability and licensing before requesting a real dataset:

```{r available-benchmarks}
benchmark_status <- available_spatial_benchmarks()
benchmark_status[c(
  "dataset", "included", "observations", "classes", "scenarios"
)]
```

DLPFC and CRC are bundled as coordinate-and-label derivatives. Expression
matrices and tissue images are excluded. MERFISH is listed for provenance but
cannot be redistributed with the package because the derived annotation does not
have an explicit redistribution license. The complete object also records the
`license`, `source`, and `note` fields for every dataset.

## `load_spatial_benchmark()`

Load a bundled scenario by integer position:

```{r load-dlpfc}
dlpfc_case <- load_spatial_benchmark("dlpfc", scenario = 1L)
c(
  observations = nrow(dlpfc_case$xy),
  classes = nlevels(dlpfc_case$truth),
  specimens = nlevels(dlpfc_case$samples),
  input_accuracy = mean(dlpfc_case$labels == dlpfc_case$truth)
)
dlpfc_case$scenario
```

Scenario identifiers are also accepted:

```{r load-crc, eval=FALSE}
crc_case <- load_spatial_benchmark(
  "crc", scenario = "CRC_random_25_r1"
)
crc_refined <- refine_spatial_labels(
  crc_case$xy, crc_case$labels,
  samples = crc_case$samples,
  workers = 4L
)
evaluate_spatial_refinement(
  crc_case$truth, crc_case$labels, crc_refined,
  boundary = crc_case$boundary,
  sparse = crc_case$sparse,
  method = "FiberMargin"
)
```

The legacy dataset name `"colorectal"` is accepted as an alias for `"crc"`.
Calling `load_spatial_benchmark("merfish")` intentionally reports why that
dataset is not bundled.

# End-to-end analysis template

The following pattern can be adapted to a real coordinate-indexed annotation:

```{r end-to-end-template, eval=FALSE}
# One row per spatial location.
xy <- as.matrix(my_data[, c("x", "y")])
labels <- factor(my_data$annotation)
samples <- factor(my_data$section_id)

# Repair does not receive the reference labels.
refined <- refine_spatial_labels(
  xy = xy,
  labels = labels,
  samples = samples,
  workers = 4L
)

# Inspect edits before replacing an annotation column.
audit <- data.frame(
  observed = labels,
  candidate = attr(refined, "candidate"),
  refined = refined,
  changed = attr(refined, "changed"),
  margin_score = attr(refined, "margin_score"),
  required = attr(refined, "required"),
  repair_margin = attr(refined, "repair_margin"),
  isolation = attr(refined, "isolation")
)

my_data$fibermargin_annotation <- refined
```

When an independent reference is available, evaluate correction and damage
together:

```{r end-to-end-evaluation, eval=FALSE}
evaluate_spatial_refinement(
  truth = reference_labels,
  initial = labels,
  refined = refined,
  boundary = boundary_locations,
  sparse = rare_or_sparse_locations,
  method = "FiberMargin"
)
```

# Reporting checklist

For reproducible analyses, report:

1. Coordinate dimension and number of locations.
2. Number of classes and independent specimens.
3. Whether the binary specialization or multiclass enclosure mechanism was used.
4. The `workers` budget and package version.
5. Accuracy and ARI when a reference exists.
6. Correction recall and damage rate together.
7. Boundary, sparse-region, macro-recall, and worst-recall diagnostics when relevant.
8. The unchanged input, repaired labels, and pointwise audit attributes.

Function-level argument and return-value details remain available through
`help(package = "fibermargin")` and `?function_name`.
