---
title: "FiberMargin Reference Manual"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fibermargin_reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette summarizes the package interface for quick look-up while working in R.

```{r setup, include=FALSE}
options(width = 78)
library(fibermargin)
```

## Functions

### Refinement

- `refine_spatial_labels(xy, labels, samples = NULL, workers = NULL)`  
  Refines noisy labels on finite 2D/3D coordinates. `samples` defines fully
  independent coordinate systems, and `workers` is one deterministic total CPU
  budget.

- `clean_categorical_mask(mask, samples = NULL, workers = NULL)`  
  Cleans a categorical matrix (2D) or array (3D).

### Simulation

- `simulate_spatial_domains(...)`  
- `simulate_complex_spatial_domains(...)`  
- `simulate_spatial_clusters(...)`  
- `simulate_gradient_regions(...)`  
- `simulate_volumetric_domains(...)`

All simulators return a `spatial_refinement_benchmark` object with `xy`,
`labels`, `truth`, `samples`, and optional `boundary`/`sparse` fields.

## `refine_spatial_labels()` contract

- `xy`: a finite numeric matrix with exactly two or three columns.
- `labels`: one non-missing categorical assignment per row.
- `samples`: optional integer, character, or factor identifiers. Coordinates and
  labels never cross sample boundaries, even when coordinate values overlap.
- `workers`: `NULL` or one positive integer specifying the total CPU budget.

Constant axes are removed separately in each sample. A constant `z` therefore
uses the 2D operator, while a variable `z` uses the genuine 3D operator. Every
sample must retain at least two varying axes.

The return value is an ordinary factor with input levels and identifiers. Its
pointwise attributes are `candidate`, `margin_score`, `required`,
`repair_margin`, `atlas_dispersion`, `isolation`, and `changed`. Summary
attributes are `workers`, `dimensions_used`, `labels_changed`,
`changed_fraction`, `classes_before`, `classes_after`, `removed_classes`, and
`sample_sizes`. Class summaries are per-sample named lists. Class preservation
is not imposed; `removed_classes` reports observed classes absent after repair.

```{r contract-examples}
set.seed(8)
xy <- matrix(runif(1200), ncol = 2)
labels <- factor(ifelse(xy[, 1] < 0.5, "left", "right"))

refined_2d <- refine_spatial_labels(xy, labels, workers = 1L)
refined_flat_3d <- refine_spatial_labels(
  cbind(xy, z = 0), labels, workers = 1L
)
stopifnot(identical(refined_2d, refined_flat_3d))

volume <- simulate_volumetric_domains(
  n = 1200L, shape = "folded_layers", samples = 2L, seed = 9L
)
refined_3d <- refine_spatial_labels(
  volume$xy, volume$labels, volume$samples, workers = 2L
)
attr(refined_3d, "dimensions_used")

overlap_xy <- rbind(xy, xy)
overlap_labels <- factor(rep(as.character(labels), 2L))
specimen <- rep(c("first", "second"), each = nrow(xy))
refined_joint <- refine_spatial_labels(
  overlap_xy, overlap_labels, specimen, workers = 2L
)
attr(refined_joint, "sample_sizes")
```

### Evaluation

- `evaluate_spatial_refinement(truth, initial, refined, ...)`  
  Computes recovery, consistency, boundary, sparse-region, and damage/repair
  diagnostics.

- `evaluate_mask_cleaning(reference, initial, cleaned, ...)`  
  Computes mean IoU, boundary IoU, and damage/repair decomposition for masks.

### Benchmarking

- `benchmark_spatial_refiners(data, methods, include_initial = TRUE, seed = 1L, ...)`  
  Runs multiple methods on one or more identical benchmark inputs.

- `spatial_benchmark(xy, labels, truth, samples, ...)`  
  Validates and constructs a benchmark object.

- `available_spatial_benchmarks()`  
  Lists bundled datasets and licensing information.

- `load_spatial_benchmark(name, scenario)`  
  Loads a bundled real scenario.

## Scoring glossary

- `accuracy`: overall corrected agreement.
- `accuracy_gain`: improvement over input labeling.
- `correction_recall`: fraction of wrong labels that become correct.
- `damage_rate`: fraction of correct labels that become wrong.
- `worst_recall`: minimum class recall.
- `boundary_accuracy`: conditional accuracy on boundary-labeled spots.
- `sparse_region_accuracy`: conditional accuracy on a sparse reference region.
- `ari`: adjusted Rand index.
- `n`: number of observations.
